
### Loading in Packages and Data

```{r}
library(tidyverse)
library(sf)
library(here)
library(tmap)
library(stars)

# read in geodatabase of EJScreen data at the Census Block Group level
ejscreen <- st_read(here("eds223-hw1", "data", "ejscreen", "EJSCREEN_2023_BG_StatePct_with_AS_CNMI_GU_VI.gdb")) 

```

---

*Region:* Orange County, California
*Community:* Low % of Native English Speakers
*EJ Issue:* Hazardous Waste Proximity

---

### Preliminary Data Loading
```{r}
# filter database to just data in orange county 
orange_county <- ejscreen |> 
    filter(CNTY_NAME == "Orange County")
```

Becuase we are looking at only orange county, we will need to filter the dataset to just entries in Orange County, California (this is where I am from). The two maps that will be displayed will show the distribution of Low % Native English Speakers, and the distribution of Hazardous waste proximity. By taking a close look at the distribution of these two, we will hopefully be able to eyeball a relationship.

In the dataset, the columns names are:
- `orange_county$LINGISOPCT` to look at the percentage of limited Englisht-speaking households
- `orange_county$D2_PTSDF` to look at the index that quanitfies the hazardous waste proximity of a community (Environmental Justice Index)

### Map 1: Spatial Distribution of Low Percentage English Speakers

```{r}
# Create variable containing spatial data for block groups in Orange County
oc_shape <- st_as_sf(orange_county$Shape, crs = 4236)
tm_shape(oc_shape) +
    tm_polygons(fill="lightblue") +
    tm_crs("auto")

```

