
### Loading in Packages and Data

```{r}
library(tidyverse)
library(sf)
library(here)
library(tmap)
library(stars)
library(viridisLite)


# read in geodatabase of EJScreen data at the Census Block Group level
ejscreen <- st_read(here(
  "eds223-hw1",
  "data",
  "ejscreen",
  "EJSCREEN_2023_BG_StatePct_with_AS_CNMI_GU_VI.gdb"
))
```

---

*Region:* Orange County, California
*Community:* Low % of Native English Speakers
*EJ Issue:* Hazardous Waste Proximity

---

### Preliminary Data Loading
```{r}
# filter database to just data in orange county
orange_county <- ejscreen |>
  filter(CNTY_NAME == "Orange County") |>
  filter(ST_ABBREV == "CA")
```

Becuase we are looking at only orange county, we will need to filter the dataset to just entries in Orange County, California (this is where I am from). The two maps that will be displayed will show the distribution of Low % Native English Speakers, and the distribution of Hazardous waste proximity. By taking a close look at the distribution of these two, we will hopefully be able to eyeball a relationship.

In the dataset, the columns names are:
- `orange_county$LINGISOPCT` to look at the percentage of limited Englisht-speaking households
- `orange_county$D2_PTSDF` to look at the index that quanitfies the hazardous waste proximity of a community (Environmental Justice Index)

### Map 1: Spatial Distribution of Low Percentage English Speakers

```{r}
# Map the spatial distribution of low percentage english speakers
tm_shape(orange_county) +
  tm_fill(
    fill = 'LINGISOPCT',
    fill.legend = tm_legend(title = "Percentage of Limited English Speakers"),
    fill.scale = tm_scale(values = c('lightblue', 'darkgreen'))
  ) +
  tm_borders(
    col = 'black',
    lwd = 0.05
  ) +
  tm_scalebar(position = c("left", "bottom")) +
  tm_compass(position = c("left", "top")) +
  tm_title(
    "Spatial Distribution of Low Percentage English Speakers",
    bg = TRUE,
    frame = TRUE,
    frame.color = 'black',
    size = 1.5
  )
```

### Map 3: Spatial Distribution of Hazardous Waste Proximity

```{r}
# Map the spatial distribution of low percentage english speakers
tm_shape(orange_county) +
  tm_fill(
    fill = 'D2_PTSDF',
    fill.legend = tm_legend(title = "Hazardous Waste Proximity EJ Index"),
    fill.scale = tm_scale(values = c('lightgreen', 'darkred'))
  ) +
  tm_borders(
    col = 'black',
    lwd = 0.05
  ) +
  tm_scalebar(position = c("left", "bottom")) +
  tm_compass(position = c("left", "top")) +
  tm_title(
    "Spatial Distribution of Hazardous Waste Proximity",
    bg = TRUE,
    frame = TRUE,
    frame.color = 'black',
    size = 1.5
  )
```